---
title: "Project 1"
author: "Nermin Bibic"
date: "6/20/2021"
output:
  github_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, error = TRUE)
```

# Reading and Summarizing data from the National Hockey League’s (NHL) API 

### Packages required to run code
```{r, echo = TRUE}
library(knitr)
library(tidyverse)
library(httr)
library(jsonlite)
library(ggplot2)
library(readr)
library(DT)
```

## Functions to contact the NHL records API

The below functions return NHL records in well-formatted, parsed data frames, with options to specify the franchise of choice by both name and ID number.

Our base URLs is: https://records.nhl.com/site/api.

```{r, echo = TRUE}
baseURLRecords <- "https://records.nhl.com/site/api/"
baseURLStats <- "https://statsapi.web.nhl.com/api/v1/teams"
```

### Function returning id, firstSeasonId, lastSeasonId, and name of every team in the history of the NHL.
```{r, echo = TRUE}
franchises <- function() {
  URL <- paste0(baseURLRecords, 'franchise')
  franchiseRAW <- RCurl::getURL(URL)
  franchiseDF <- fromJSON(franchiseRAW, flatten=TRUE)
  franchiseDF <- as_tibble(franchiseDF$data)
  franchiseDF <- franchiseDF %>% select(id,
                                        mostRecentTeamId,
                                        fullName,
                                        firstSeasonId,
                                        lastSeasonId)
  return(franchiseDF)
}
```

### Function returning total stats for every franchise (ex: roadTies, roadWins, etc.)
```{r, echo = TRUE}
franchiseTeamTotals <- function() {
  URL <- paste0(baseURLRecords, 'franchise-team-totals')
  franchiseTeamTotalsRAW <- RCurl::getURL(URL)
  franchiseTeamTotalsDF <- fromJSON(franchiseTeamTotalsRAW, flatten=TRUE)
  franchiseTeamTotalsDF <- as_tibble(franchiseTeamTotalsDF$data)
  return(franchiseTeamTotalsDF)
}
```

### Function returning season records for a specific franchise
```{r, echo = TRUE}
seasonRecords <- function(team, method = c('franchise_id', 'team_id', 'team_name')) {
  if(missing(team)) {
    stop('Please specify team')
  }
  if(missing(method)) {
    stop('Please specify method. Acceptable methods:\n\nfranchise_id\nteam_id\nteam_name')
  }
  if (method == 'team_id') {
    team = select(filter(franchises(), mostRecentTeamId == team), id)$id
  } else if (method == 'team_name') {
    team = select(filter(franchises(), fullName == team), id)$id
  }
  URL <- paste0(baseURLRecords, 'franchise-season-records?cayenneExp=franchiseId=', team)
  seasonRecordsRAW <- RCurl::getURL(URL)
  seasonRecordsDF <- fromJSON(seasonRecordsRAW, flatten=TRUE)
  seasonRecordsDF <- as_tibble(seasonRecordsDF$data)
  return(seasonRecordsDF)
}
```

### Function returning season records for a specific franchise
```{r, echo = TRUE}
goalieRecords <- function(team, method = c('franchise_id', 'team_id', 'team_name')) {
  if(missing(team)) {
    stop('Please specify team')
  }
  if(missing(method)) {
    stop('Please specify method. Acceptable methods:\n\nfranchise_id\nteam_id\nteam_name')
  }
  if (method == 'team_id') {
    team = select(filter(franchises(), mostRecentTeamId == team), id)$id
  } else if (method == 'team_name') {
    team = select(filter(franchises(), fullName == team), id)$id
  }
  URL <- paste0(baseURLRecords, 'franchise-goalie-records?cayenneExp=franchiseId=', team)
  goalieRecordsRAW <- RCurl::getURL(URL)
  goalieRecordsDF <- fromJSON(goalieRecordsRAW, flatten=TRUE)
  goalieRecordsDF <- as_tibble(goalieRecordsDF$data)
  return(goalieRecordsDF)
}
```

### Function returning skater records for a specific franchise
```{r, echo = TRUE}
skaterRecords <- function(team, method = c('franchise_id', 'team_id', 'team_name')) {
  if(missing(team)) {
    stop('Please specify team')
  }
  if(missing(method)) {
    stop('Please specify method. Acceptable methods:\n\nfranchise_id\nteam_id\nteam_name')
  }
  if (method == 'team_id') {
    team = select(filter(franchises(), mostRecentTeamId == team), id)$id
  } else if (method == 'team_name') {
    team = select(filter(franchises(), fullName == team), id)$id
  }
  URL <- paste0(baseURLRecords, 'franchise-skater-records?cayenneExp=franchiseId=', team)
  skaterRecordsRAW <- RCurl::getURL(URL)
  skaterRecordsDF <- fromJSON(skaterRecordsRAW, flatten=TRUE)
  skaterRecordsDF <- as_tibble(skaterRecordsDF$data)
  return(skaterRecordsDF)
}
```

### Function returning admininstration history and retired numbers
```{r, echo = TRUE}
adminHistory <- function(team, method = c('franchise_id', 'team_id', 'team_name')) {
  if(missing(team)) {
    stop('Please specify team')
  }
  if(missing(method)) {
    stop('Please specify method. Acceptable methods:\n\nfranchise_id\nteam_id\nteam_name')
  }
  if (method == 'franchise_id') {
    team = select(filter(franchises(), id == team), mostRecentTeamId)$mostRecentTeamId
  } else if (method == 'team_name') {
    team = select(filter(franchises(), fullName == team), mostRecentTeamId)$mostRecentTeamId
  }
  URL <- paste0(baseURLRecords, 'franchise-detail?cayenneExp=mostRecentTeamId=', team)
  adminHistoryRAW <- RCurl::getURL(URL)
  adminHistoryDF <- fromJSON(adminHistoryRAW, flatten=TRUE)
  adminHistoryDF <- as_tibble(adminHistoryDF$data)
  return(adminHistoryDF)
}
```

## Function to contact the NHL stats API for the ?expand=team.stats modifier.

This function returns team stats. If no team is specified (by running the function as "teamStats()"), this function returns stats for all teams. If a team is specified, either by francise ID, team ID, or team name, this function returns a single row of stats data. The stats data contain both raw values for each data point as well as ranks for the data points.

```{r, echo = TRUE}
teamStats <- function(team, method = c('franchise_id', 'team_id', 'team_name')) {
  
  if (missing(team)) {
    
    URL <- paste0(baseURLStats, '?expand=team.stats')
    teamStatsRAW <- RCurl::getURL(URL)
    teamStatsDF <- fromJSON(teamStatsRAW, flatten=TRUE)
    teamStatsList <- teamStatsDF$teams$teamStats
    # Remove NULL objects
    teamStatsList <- teamStatsList[-which(sapply(teamStatsList, is.null))]
    # Initialize empty list, to be filled with single-row tibbles
    allTeamsStats <- list()
    for (i in teamStatsList) {
      teamStatsSplits <- as_tibble(as.data.frame(i$splits))
      # Split raw value and rank data
      teamStatsValues <- teamStatsSplits[1, ]
      teamStatsRanks <- select(teamStatsSplits[2, ], -c('team.id', 'team.name', 'team.link'))
      # Clean up column names; remove prefix
      colnames(teamStatsValues) <- gsub("stat.", "", colnames(teamStatsValues))
      colnames(teamStatsRanks) <- gsub("Rank", "", colnames(teamStatsRanks))
      colnames(teamStatsRanks) <- gsub("stat.", "", colnames(teamStatsRanks))
      # Distinguish rank data
      colnames(teamStatsRanks) <- paste0(colnames(teamStatsRanks), 'Rank')
      # Drop NA columns
      teamStatsValues <- teamStatsValues[, colSums(is.na(teamStatsValues)) != nrow(teamStatsValues)]
      teamStatsRanks <- teamStatsRanks[, colSums(is.na(teamStatsRanks)) != nrow(teamStatsRanks)]
      # Combine values and ranks into one row
      teamStatsBind <- as_tibble(cbind(teamStatsValues, teamStatsRanks))
      allTeamsStats[[length(allTeamsStats) + 1]] <- teamStatsBind
    }
    # combine all tibbles in resulting list into one tibble
    allTeamsStats <- bind_rows(allTeamsStats)
    return(allTeamsStats)

  } else {
    
    if(missing(team)) {
      stop('Please specify team')
    }
    if(missing(method)) {
      stop('Please specify method. Acceptable methods:\n\nfranchise_id\nteam_id\nteam_name')
    }
    if (method == 'franchise_id') {
      team = select(filter(franchises(), id == team), mostRecentTeamId)$mostRecentTeamId
    } else if (method == 'team_name') {
      team = select(filter(franchises(), fullName == team), mostRecentTeamId)$mostRecentTeamId
    }
    URL <- paste0(baseURLStats, '/', team, '?expand=team.stats')
    teamStatsRAW <- RCurl::getURL(URL)
    teamStatsDF <- fromJSON(teamStatsRAW, flatten=TRUE)
    # Choose 'teamStats' object in JSON
    teamStatsDF <- as_tibble(teamStatsDF$teams$teamStats[[1]], .name_repair = 'minimal')
    teamStatsSplits <- as_tibble(as.data.frame(teamStatsDF$splits))
    # Split raw value and rank data
    teamStatsValues <- teamStatsSplits[1, ]
    teamStatsRanks <- select(teamStatsSplits[2, ], -c('team.id', 'team.name', 'team.link'))
    # Clean up column names; remove prefix
    colnames(teamStatsValues) <- gsub("stat.", "", colnames(teamStatsValues))
    colnames(teamStatsRanks) <- gsub("Rank", "", colnames(teamStatsRanks))
    colnames(teamStatsRanks) <- gsub("stat.", "", colnames(teamStatsRanks))
    # Distinguish rank data
    colnames(teamStatsRanks) <- paste0(colnames(teamStatsRanks), 'Rank')
    # Drop NA columns
    teamStatsValues <- teamStatsValues[, colSums(is.na(teamStatsValues)) != nrow(teamStatsValues)]
    teamStatsRanks <- teamStatsRanks[, colSums(is.na(teamStatsRanks)) != nrow(teamStatsRanks)]
    # Combine values and ranks into one row
    teamStatsBind <- as_tibble(cbind(teamStatsValues, teamStatsRanks))
    return(teamStatsBind)
    
  }
  
}
```

## Wrapper function
This wrapper function is a one-stop-shop for the user to access any of the API endpoints above. This function simply calls the appropriate endpoint as per the users request, including any modifiers, team IDs, etc.

```{r, echo = TRUE}
NHLData <- function(target_data = c('franchises', 'team_totals', 'season_records',
                                    'goalie_records', 'skater_records',
                                    'admin_history', 'team_stats'),...) {
  switch(target_data,
           franchises = franchises(),
           team_totals = franchiseTeamTotals(),
           season_records = seasonRecords(...),
           goalie_records = goalieRecords(...),
           skater_records = skaterRecords(...),
           admin_history = adminHistory(...),
           team_stats = teamStats(...),
           stop("Please use valid argument located in function definition line.")
  )
}
```

## Exploratory Data Analysis

```{r, echo = TRUE}
rangersSkaterRecords <- skaterRecords(3, 'team_id')
rangersGoalieRecords <- goalieRecords(3, 'team_id')
# Combine skaters and goalies regardless of different sets of columns
rangersPlayerRecords <- bind_rows(rangersSkaterRecords, rangersGoalieRecords)
```

### Contingency Tables

Here is a two-way contingency table between position and games played showing the frequency of each category.

```{r, echo = TRUE}
rangersPlayerRecords$gamesPlayedRange <- cut(rangersPlayerRecords$gamesPlayed,
                                             c(0, 100, 200, 300, 400, 500,
                                               max(rangersPlayerRecords$gamesPlayed)))
levels(rangersPlayerRecords$gamesPlayedRange) = c("<100", "101-200", "201-300",
                                                  "301-400", "401-500", ">500")

rangersPlayerRecords$position <- as.factor(rangersPlayerRecords$positionCode)
levels(rangersPlayerRecords$position) <- list("Goalie" = "G",
                                              "Defender" = "D",
                                              "Right Winger" = "R",
                                              "Left Winger" = "L",
                                              "Center" = "C")

twoWayTab <- table(rangersPlayerRecords$position,
                   rangersPlayerRecords$gamesPlayedRange)
kable(twoWayTab, caption = 'Position and Games Played')
```

Here is a three-way contingency table. We are obtaining bivariate info form this three-dimensional array. First we are showing a two-way contingency table between position and games played for inactive players, and then a two-way contingency table between position and games played for active players.

```{r, echo = TRUE}
threeWayTab <- table(rangersPlayerRecords$position,
                     rangersPlayerRecords$gamesPlayedRange,
                     rangersPlayerRecords$activePlayer)
kable(threeWayTab[ , ,1], caption = "For inactive players:")
kable(threeWayTab[ , ,2], caption = "For active players:")
```

Here is a two-way contingency table between position and goals/assists per game. I dropped all players who played 50 or less games so the goal/assist per game proficiency is more accurate (a player who played 1 game and scored 1 goal would have a rate of 1 goal/assist per game and would appear to be the most proficient scorer; for this reason I am dropping players who didn't play many games).

```{r, echo = TRUE}
rangersPlayerRecords$goalOrAssistPerGame <- (rangersPlayerRecords$goals +
                                             rangersPlayerRecords$assists) /
                                            rangersPlayerRecords$gamesPlayed

rangersPlayerRecords <- droplevels(rangersPlayerRecords)

# Count only outfield players who played over 50 games, and convert NA values to 0
rangersPlayerRecords <- rangersPlayerRecords %>% filter(gamesPlayed > 50,
                                                        !is.na(goalOrAssistPerGame),
                                                        positionCode != 'G')

# Create goals/assist per game ranges
rangersPlayerRecords$GAPerGameRange <- cut(rangersPlayerRecords$goalOrAssistPerGame,
                                           c(0, .2, .4, .6, .8, 1,
                                             max(rangersPlayerRecords$goalOrAssistPerGame)))
levels(rangersPlayerRecords$GAPerGameRange) = c("<0.2", "0.2-0.4", "0.4-0.6",
                                                  "0.6-0.8", "0.8-1", ">1")

twoWayTab <- table(rangersPlayerRecords$position,
                   rangersPlayerRecords$GAPerGameRange)
kable(twoWayTab, caption = "Position and Goal/Assist Per Game")
```

### Numerical Summaries of Outfield Player Performance Data
The summary statistics for each of the variables are grouped below separately, by the outfield positions.

I created two new variables for goals per game and assists per game.

```{r, echo = TRUE}
rangersPlayerRecords$goalsPerGame <- rangersPlayerRecords$goals /
                                     rangersPlayerRecords$gamesPlayed
rangersPlayerRecords$assistsPerGame <- rangersPlayerRecords$assists /
                                       rangersPlayerRecords$gamesPlayed

# Subset by columns we want to analyze
rangersPlayerStats <- rangersPlayerRecords[ , c("gamesPlayed",
                                                "goalsPerGame",
                                                "assistsPerGame",
                                                "goalOrAssistPerGame",
                                                "position")]

# Function for summary statistics for the numerical columns
summarizePositions <- function(df, pos) {
  dfFiltered <- filter(df, position == pos)[, -5]
  kable(do.call(cbind, lapply(dfFiltered, summary, digits = 3)),
        caption = paste("Summary of Position", pos))
}
summarizePositions(rangersPlayerStats, "Defender")
summarizePositions(rangersPlayerStats, "Center")
summarizePositions(rangersPlayerStats, "Right Winger")
summarizePositions(rangersPlayerStats, "Left Winger")
```

Looking at the mean and medians for the above summary data, it appears skaters on the right wing are more effective at scoring and assisting goals than those on the left wing. As expected, centers (the main attacking position) have the highest mean and median goal or assist per game.

### Barplot of Goals or Assists Per Game Range
Each of the below bar graphs show counts based on two variables. These bar graphs are side-by-side bar plots, allowing us to compare counts between more than one factor.

explain
```{r, echo = TRUE}
g <- ggplot(rangersPlayerRecords, aes(x = GAPerGameRange))
g + geom_bar(aes(fill = position), position = "dodge") +
    labs(x = "Goals/Assists Per Game Ranges") +
    scale_fill_discrete(name = "Skater Position")
```

### Kernel Smoother Histogram for Goals or Assists Per Game
explain

```{r, echo = TRUE}
g <- ggplot(rangersPlayerRecords, aes(x = goalOrAssistPerGame))
g + geom_histogram(aes(y = ..density..), binwidth = 0.05) +
    geom_density(alpha = 0.5, position = "stack", color="green", lwd = 2) +
    labs(title = "Histogram for Goals/Assists Per Game Range", y = "Density") +
    facet_wrap(~ position)
```

### Scatterplot of Games Played and Penalty Minutes
This scatter plot shows the relationship between games played and penalty minutes. This plot is grouped by player position. Lines have been added to fit the dots on the graph using the linear model method. For all positions as the player plays more games, they steadily amass penalty minutes. This is mostly true for defenders who appear to gain penalty minutes at a greater rate the more they play. Surprisingly, the rate of increase is more similar between right wingers and centers than between right wingers and left wingers.

```{r, echo = TRUE}
rangersPlayerRecordsU500 <- rangersPlayerRecords %>% filter(gamesPlayed < 500)
g <- ggplot(rangersPlayerRecordsU500, aes(x = gamesPlayed, y = penaltyMinutes, group = position))
g + geom_point(aes(col = position), size = 1) +
    geom_smooth(method = lm, aes(col = position)) +
    labs(title = "Penalty Minutes vs. Points")
```

### Boxplot for Games Played by Position
The box plot below measures the spread of the number of games played for players by position. The few outliers with over 500 games have been filtered out to reduce the scale of the y-axis in order to better discern the difference among the largest crop of players. The number of games played among all positions is very similar. There is a lower spread among goalkeepers and there are far less goalkeepers overall. There are very few goalkeepers and a signficant amount of outfield players with more than 150 games to their name.

```{r, echo = TRUE}
rangersPlayerRecords <- bind_rows(rangersSkaterRecords, rangersGoalieRecords)
rangersPlayerRecordsU500 <- rangersPlayerRecords %>% filter(gamesPlayed < 500)
g <- ggplot(rangersPlayerRecordsU500, aes(x = positionCode, y = gamesPlayed))
g + geom_boxplot() +
    geom_point(aes(col = positionCode), alpha = 1, size = 0.9, position = "jitter") +
    labs(title = "Boxplot for Games Played", x = "Position", y = "Games Played")
```

### Boxplot for Goals Per Game
Among very experienced players who have played over 150 games, there appears to be a slightly larger spread in the goals per game rate among right wing players than for left wing and center players. The data is more spread out for the attacking positions. There are many wing and center players in the lower tails who score as many per game as defenders. The spread in the goals per game rate among experienced defenders is much smaller than for the other positions, mainly situated between 0.05 and 0.1 goals per game.

```{r, echo = TRUE}
rangersPlayerRecords$goalsPerGame <- rangersPlayerRecords$goals /
                                     rangersPlayerRecords$gamesPlayed
rangersPlayerRecords <- rangersPlayerRecords %>% filter(gamesPlayed > 150,
                                                        !is.na(goalsPerGame),
                                                        positionCode != 'G')
g <- ggplot(rangersPlayerRecords, aes(x = positionCode, y = goalsPerGame))
g + geom_boxplot() +
    geom_point(aes(col = positionCode), alpha = 1, size = 0.9, position = "jitter") +
    labs(title = "Boxplot for Goals Per Game", x = "Position", y = "Goals Per Game")
```




One more plot?
