---
title: "Project 1"
author: "Nermin Bibic"
date: "6/20/2021"
output:
  github_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, error = TRUE)
```

# Reading and Summarizing data from the National Hockey League’s (NHL) API 

### Packages required to run code
```{r, echo = TRUE}
library(knitr)
library(tidyverse)
library(ggplot2)
library(readr)
library(DT)
```

### Functions 

•	You should write functions to contact the NHL records API for the endpoints listed below. The functions should return well-formatted, parsed data (usually a data frame). Where possible, the user
should have the option to specify the franchise of choice by both name and ID number - you’ll need to map the names to ID numbers yourself.
–	/franchise (Returns id, firstSeasonId and lastSeasonId and name of every team in the history of the NHL)
–	/franchise-team-totals (Returns Total stats for every franchise (ex roadTies, roadWins, etc))
–	/site/api/franchise-season-records?cayenneExp=franchiseId=ID (Drill-down into season records for a specific franchise)
–	/franchise-goalie-records?cayenneExp=franchiseId=ID (Goalie records for the specified franchise)
–	/franchise-skater-records?cayenneExp=franchiseId=ID (Skater records, same interaction as goalie endpoint)
–	/site/api/franchise-detail?cayenneExp=mostRecentTeamId=ID (Admin history and retired num- bers)
